### 2022-02-18 paper/paper.tex
adding paper, finishing model in rough state

\documentclass{article}
\usepackage{amsmath}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=blue,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
}

\begin{document}

\title{Do professional chess players draw too often?}
\author{Tudor Muntianu}
\maketitle

\section*{Disclaimer}
I spent most of my time trying to get the models working. It was a complete nightmare. Getting the factor analysis model to work (replacing Smerdon's PCA) took me about 10 hours.
Maybe I should have settled for PCA for this draft.
Perhaps I should not have tried to relearn Pyro, a probabilistic programming language written in Python.
That being said, if I had to build this model in R I likely would have ended up using RStan, which is a binding to another PPL, so the model wouldn't have really been in R either.
The data cleaning and manipulation would also have taken me 10x longer since I don't really know R.
Hindsight is 20/20.

The code (and this writeup) is in a rough state as a result -- by the final draft it will be pretty and commented and clear (despite being in Python).
Similarly, since the model was giving me issues I didn't have compute time to train the model on the whole dataset (which is 4.6M games).
Again this will not be the case by the final draft. Everything works though; I have a complete pipeline
that just needs some clean up, some hyperparameters to be adjusted/played with, and to be run to completion.

Also, there are a couple of things I would like to add (or at least investigate) by the final draft:
\begin{itemize}
    \item Dataset of online games (ostensibly less incentive to draw here, more ``naturalistic'' or aggressive gameplay)
    \item Using the factor model to compute an index, comparing to Smerdon's results
\end{itemize}

All code can be found \href{https://github.com/tmuntianu/chessdraws}{here}.

\section{Introduction}

The prevalence of draws in professional chess has been steadily increasing over the past couple decades.
As play becomes more precise and accurate due to extensive computer preparation and analysis,
and as more money enters the sport, winning edges are diminished and players are increasingly
incentivized not to lose.
For instance, at the 2018 World Chess Championship, where Magnus Carlsen played Fabiano Caruana (World No. 1 and No.2 respectively),
all 12 classical time-control games were drawn.
The championship was decided by a rapid chess tiebreaker.
Near-perfect play is not the only factor, however.
Since tournament fates, prize pools, and sponsorship deals hinge on performance, losing is highly undesirable.
So, the popular argument goes, players play for and agree upon too many draws. This opinion is shared by
many strong players and professionals, notably Susan Polgar, a very strong grandmaster who argues that agreed draws should be eliminated from the game.
While some agreed draws are legitimate reflections of drawn game state, others occur when players simply play out opening moves and subsequently agree to draw
to minimize variance. This sort of behavior is most egregious, but it demonstrates the most extreme form of this behavior, making it likely
that playing passively or ``soft-playing'' for draws also occurs.

The question we want to answer is: are draws in high-level chess a result of optimal play?
or are players sacrificing chances to win to reduce variance as a result of external pressures?
Playing ``solid'' chess (i.e., slow, methodical, less aggressive) that lends itself to more draws is not what we are investigating,
but rather if players are playing worse as a result of draw incentives. In other words, are players
playing below their true strengths as a result of playing for draw?
Put even another way, are players are playing to maximize the single-game outcome (i.e., win if possible, draw if losing),
or are players sacrificing performance to reduce their variance?
% From here out, irrational play will be used to refer to this playstyle. This is not technically true, since the payoffs for players are not just winning or losing, but reputation, sponsorships, tournament placement, etc.
% But for the purposes of this paper, playing rationally means playing to maximize the outcome of an individual game, not playing to maximize the real-world consequences.

In this paper the frequency of draws is estimated and compared to existing estimates, namely the very robust Elo rating system used in professional play. We begin by adapting David Smerdon's Fighting Chess Index to generate latent variables representing player behavior and aggression,
then use those latent variables in a simple Bayesian model to estimate the probability of winning, drawing, and losing.
Then, using the Elo rating system, we compute the expected value of the game's outcome, and compare our estimated probabilities to those generated by ratings to determine whether drawing behavior results in poorer results.

\section{The game}

In this paper we make a number of assumptions about each chess game observed.
First, we assume that each player plays at the strength of their respective Elo rating, or
in other words, plays their best. We make no attempt to model player strength directly or
measure the variance in their performance.
Instead, we want to know how aggressive or how passive they are playing.

Put more formally, given two players $W$ and $B$, we want to know
$p=(p_{\text{win}}, p_{\text{draw}}, p_{\text{lose}})$, where $p_{\text{outcome}}$ is
the probability of $W$ (white) winning, drawing, and losing respectively.

The Elo system, while it does not give a perfect estimate of $p$, does shed some light.

\subsection{Elo ratings}
The Elo rating system gives an excellent estimate of the strength of a player.
With sufficient games played against many opponents, its application to chess produces robust results. All major chess governing bodies (USCF and FIDE) use Elo as their rating system,
chess engines are benchmarked using Elo, and it is used ubiquitously in tournaments to determine seeding and matches.

Given the ratings $R_A$ and $R_B$ of two players $A$ and $B$ respectively, the Elo rating system functions as follows.
\begin{align*}
    E_A = \frac{1}{1+10^{(R_B-R_A)/400}} \\
    \text{where } E_A = p_{\text{win}} + \frac{1}{2} p_{\text{draw}}
\end{align*}
It follows from this that $E_A+E_B=1$, unsurprisingly,

\subsection{Player strategies and equilibrium}

The Elo rating system makes an implicit assumption that draw frequency does not directly affect
the expected score $E$.
There are infinite solutions to $E = p_{\text{win}} + \frac{1}{2} p_{\text{draw}}$.
So, to simplify chess for our purposes, we assume that each player plays
according to their rating, and the choice they must make is how aggressive or passive to play.

If we assume, for instance, that both players $W$ and $B$, are playing at maximal aggression, then $E_W=p_{\text{win}}$ and $E_B=p_{\text{lose}}$. That is, $p_{\text{draw}}=0$.
But if both players are not maximally aggressive, the probability of a draw increases.
To write out all the strategies, we could define an index $a_W, a_B$ for each player that indicates their aggression,
and then rebalance
$p=(p_{\text{win}}, p_{\text{draw}}, p_{\text{lose}})$ accordingly, but enumerating the strategies formally turns out not to matter.
This is because, under Elo assumptions, $E_W$ and $E_B$ are constant, regardless of draw probability.
So while players can change the probability of drawing depending on their strategies, they
cannot change their expected score.
% below is not correct, fix the math
% Both players can contribute to this probability, so $p_{\text{draw}}=p_{\text{draw}_W}+p_{\text{draw}_B}$,
% where $p_{\text{draw}_W}$ represents $W$'s passive play or propensity to draw.
% Under Elo assumptions, without changing their expected score, a player $A$ can choose to reduce
% $p_{\text{win}_A}$ by $f$ to increase $p_{\text{draw}_A}$ by $2f$ (and therefore increase $p_{\text{draw}}$ as well).
%
% Because this adjustment can result in $p_{\text{draw}}>1$, we limit each player's contribution to $0.5$, that is
% $p_{\text{draw}_A}<0.5$.
% This allows the total probability of a draw $p_{\text{draw}}$ to be $1$, which while unrealistic

This makes any set of strategies chosen by both players a Nash equilibrium under Elo assumptions, since
the expected score for any two strategies is equal.
Intuitively, under Elo's system, players can play well
aggressively and they can play well passively; the choice of playstyle will not affect the outcome.

Our models will estimate these probabilities $p$ in order to determine whether these draw probabilities are in accordance with Elo's system, that is, reflective of player ability and playstyle,
or if they result from ``irrational play.''

The prediction, as described by Polgar \href{https://chessdailynews.com/the-700-lbs-gorilla-issue-to-draw-or-not-to-draw/}{here}, is that real-world circumstances encourage drawing at the expense of good chess.
As a result of chess being played sub-optimally, we would expect that as a result of a player playing for a draw more aggressively,
they sacrifice some results. In other words, if you play to minimize variance, you decrease your expected score.

\section{The models}
The dataset used was Caissabase, a free database of over 5M professional chess games.
Simul games, Chess960 games, and consultation (team) games were removed.
Games played by players with less than 10 games were also removed.

(I will make better citations and give more background and explanations of these models by the final draft I promise. I am rushing at the moment.)

\subsection{Player model}
For modelling player behavior, a factor model was used, inspired by David Smerdon's
\href{https://www.davidsmerdon.com/?p=2168}{Fighting Chess Index}.
Variables computed for each player were
\begin{itemize}
    \item total games
    \item average rating
    \item average absolute difference between player ratings
    \item average number of draws
    \item average number of draws in 30 moves or less
    \item average number of draws in 30 moves or less when player was White
    \item average number of moves in drawn games
\end{itemize}
See Section 12.2.4 in Bishop's \textit{Pattern Recognition and Machine Learning} for a detailed description of this factor analysis model.

Smerdon uses PCA to form his Fighting Chess Index, but this method has its flaws. As described in Bishop,
PCA can be formulated as a maximum likelihood solution of a latent variable model.
But PCA does not treat the observed variables as conditionally independent given the latent variables.
So if we use principal components to measure ``aggression'' or ``fighting spirit,'' we are not treating
average rating (a measure of player strength) and irrational draw behavior as independent.
This is problem since they are correlated.
More serious (stronger) players have more on the line, and so will generally play for draws more often.
But as we want to measure ``fighting spirit'' for individual player independent of their strength, we do not want this dependence.

In Smerdon's defense, PCA's components do lend themselves to the construction of his index (whereas the factor model used here does not), but this does not change the problem of the underlying assumption.

The factor model used here does treat observations as conditionally independent of the latent variables.
This is tremendously useful, since we want to retain measures of player strength while independently encoding
draw behavior, especially as inputs to the following model.

\subsection{Game model}
(Again I will provide more detail wrt inference, etc in final, in a rush to the deadline as a result of repeated disasters)

For each game, we modelled $p=(p_{\text{win}}, p_{\text{draw}}, p_{\text{lose}})$
using the following generative model, where $l_w, l_b$ are the latent variables
 computed from the player model for the corresponding players.
\begin{align*}
    p = A l_w + B l_b + \epsilon \\
    \epsilon \sim \mathcal{N}(\mathbf{0}, \sigma\mathbf{I}) \\
    \sigma\sim \Gamma(\alpha=1,\beta=2)
\end{align*}

\section{Results}

\end{document}